---
title: "Visualizations"
author: "Danny Morris"
date: "3/22/2022"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: false
    number_sections: true
    theme: cosmo
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

# Packages

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(viridisLite)
library(reactable)
library(plotly)
library(htmltools)
library(htmlwidgets)
```

# Data

```{r}
churn <- read_csv("churn.csv")
insurance <- read_csv("insurance.csv")
```

```{r}
insurance_sample <- insurance %>%
  group_by(smoker) %>%
  sample_n(5) %>%
  ungroup() %>%
  mutate(Id = row_number()) %>%
  select(Id, smoker, sex, age, charges)

churn_sample <- churn %>%
  group_by(churn) %>%
  slice(1:5) %>%
  ungroup() %>%
  mutate(Id = row_number()) %>%
  select(Id, number_customer_service_calls, total_night_charge, churn) 
```

# Key concepts

## Classification

```{r}
churn_sample %>%
  reactable(
    compact = T,
    bordered = TRUE, 
    columns = list(
      Id = colDef(
        name = "Customer ID",
        align = "left"
      ),
      number_customer_service_calls = colDef(
        name = "Customer Service Calls",
        align = "left"
      ),
      total_night_charge = colDef(
        name = "Daytime charges",
        align = "left"
      ),
      churn = colDef(
        name = "Churn",
        style = function(value) {
          list(
            background = "#e8e8e8"
          )
        }
      )
    )
  ) %>%
  prependContent(
    .,
    h3(class = "title", "Example data for classifying whether or not customers will churn")
  )
```

## Regression

```{r}
insurance_sample %>%
  reactable(
    compact = T,
    bordered = TRUE, 
    columns = list(
      Id = colDef(
        name = "Customer ID",
        align = "left"
      ),
      smoker = colDef(
        name = "Smoker",
        align = "left"
      ),
      sex = colDef(
        name = "Gender",
        align = "left"
      ),
      age = colDef(
        name = "Age",
        align = "left"
      ),
      charges = colDef(
        name = "Healthcare Costs",
        align = "left",
        format = colFormat(digits = 2, separators = T),
        style = function(value) {
          list(background = "#e8e8e8")
        }
      )
    )
  ) %>%
  prependContent(
    .,
    h3(class = "title", "Example data for predicting individual healthcare costs")
  )

```

## Target variable and features

```{r}
insurance_sample %>%
  reactable(
    compact = T,
    bordered = FALSE, 
    columns = list(
      Id = colDef(
        name = "Customer ID",
        align = "left",
        headerStyle = list(
          color = "#b2b2b2"
        )
      ),
      smoker = colDef(
        name = "Smoker",
        align = "left",
        headerStyle = list(
          color = "#b2b2b2"
        )
      ),
      sex = colDef(
        name = "Gender",
        align = "left",
        headerStyle = list(
          color = "#b2b2b2"
        )
      ),
      age = colDef(
        name = "Age",
        align = "left",
        headerStyle = list(
          color = "#b2b2b2"
        )
      ),
      charges = colDef(
        name = "Healthcare Costs",
        align = "left",
        headerStyle = list(
          color = "#b2b2b2"
        ),
        format = colFormat(
          digits = 2, 
          separators = T
        )
      )
    ),
    columnGroups = list(
      colGroup(
        name = "Features", 
        columns = c("age", "sex", "smoker"),
        headerStyle = list(
          color = "#0FA3B1"
        )
      ),
      colGroup(
        name = "Target", 
        columns = "charges",
        headerStyle = list(
          color = "#0FA3B1"
        )
      )
    )
  ) %>%
  prependContent(
    .,
    h3(class = "title", "Features and target variable")
  )
```

## Instances

```{r}

```

## Training, validation, testing

```{r}
insurance_sample %>%
  select(-sex) %>%
  mutate(Partition = c(rep("Training", 6), rep("Validation", 2), rep("Testing", 2))) %>%
  reactable(
    compact = T,
    bordered = TRUE, 
    columns = list(
      Partition = colDef(
        style = function(value) {
          color <- if (value == "Training") {
            "#0FA3B199"
          } else if (value == "Validation") {
            "#f0824899"
          } else {
            "#74cdfc99"
          }
          list(background = color)
        }
      ),
      Id = colDef(
        name = "Customer ID",
        align = "left"
      ),
      smoker = colDef(
        name = "Smoker",
        align = "left"
      ),
      age = colDef(
        name = "Age",
        align = "left"
      ),
      charges = colDef(
        name = "Healthcare Costs",
        align = "left",
        format = colFormat(
          digits = 2, 
          separators = T
        )
      )
    )
  ) %>%
  prependContent(
    .,
    h3(class = "title", "Example data split into training, validation, and testing partitions")
  )
```

## Actual v. Predicted

```{r}
tibble(
  Dummy = c("Actual churn: Yes", "Actual churn: No"),
  Predicted_churn = c(534, 29),
  Predicted_churn_no = c(78, 452)
) %>%
  reactable(
    compact = T,
    bordered = T,
    columns = list(
      Dummy = colDef(
        name = "",
        style = function(value) {
          list(
            fontWeight = 600
          )
        }
      ),
      Predicted_churn = colDef(
        name = "Predicted churn: Yes"
      ),
      Predicted_churn_no = colDef(
        name = "Predicted churn: No"
      )
    )
  ) %>%
  prependContent(
    x = .,
    h3(class = "title", "Example confusion matrix for comparing actual and predicted classifications")
  )
```

## Missing values

```{r}
churn_sample2 <- as.data.frame(churn_sample[,-3])
churn_sample2[c(1,7), 2] <- "NA"
churn_sample2[c(3), 3] <- "NA"

reactable(
  data = churn_sample2,
  compact = T,
  bordered = T,
  columns = list(
    Id = colDef(
      name = "Customer ID",
      align = "left"
    ),
    number_customer_service_calls = colDef(
      name = "Customer Service Calls",
      style = function(value) {
        color <- if (value == "NA") {
          "#e8e8e8"
        }
        list(background = color)
      }
    ),
    churn = colDef(
      name = "Churn",
      style = function(value) {
        color <- if (value == "NA") {
          "#e8e8e8"
        }
        list(background = color)
      }
    )
  )
) %>%
  prependContent(
    x = .,
    h3(class = "title", "Example data containing missing values")
  )
```

## Feature importance

```{r}
tibble(
  Feature = c("Smoker", "Age", "Body Mass Index"),
  Importance = c(0.87, 0.45, 0.2)
) %>%
  reactable(
    compact = T,
    bordered = TRUE
  ) %>%
  prependContent(
    .,
    h3(class = "title", "Example feature importances sorted from most to least important")
  )
```

## Overfitting

```{r}
tibble(
  Partition = c("Training", "Validation", "Testing"),
  Accuracy = c(0.93, 0.78, 0.48)
) %>%
  reactable(
    compact = T,
    bordered = TRUE
  ) %>%
  prependContent(
    .,
    h3(class = "title", "Example of overfitting where high accuracy is achieved on the training data but not the testing data.")
  )
```

## Cross validation

### Example of cross validation where the original data is partitioned several times for testing. {-}

```{r, echo=F}
browsable(
  tagList(
    list(
      tags$div(
        style = 'width:33%;display:block;float:left;padding-right:10px;border:1px solid #e8e8e8',
        reactable(
          data = insurance_sample %>% 
            select(Id) %>%
            mutate(Partition = c(rep("Training", 7), rep("Testing", 3))),
          compact = T,
          bordered = F,
          columns = list(
            Id = colDef(
              name = "Customer ID",
              align = "left"
            ),
            Partition = colDef(
              style = function(value) {
                color <- if (value == "Training") {
                  "#0FA3B199"
                } else {
                  "#f0824899"
                }
                list(background = color)
              }
            )
          )
        )
      ),
      tags$div(
        style = 'width:33%;display:block;float:left;padding-right:10px;border:1px solid #e8e8e8',
        reactable(
          data = insurance_sample %>% 
            select(Id) %>%
            mutate(Partition = c(rep("Training", 3), rep("Testing", 3), rep("Training", 4))),
          compact = T,
          bordered = F,
          columns = list(
            Id = colDef(
              name = "Customer ID",
              align = "left"
            ),
            Partition = colDef(
              style = function(value) {
                color <- if (value == "Training") {
                  "#0FA3B199"
                } else {
                  "#f0824899"
                }
                list(background = color)
              }
            )
          )
        )
      ),
      tags$div(
        style = 'width:33%;display:block;float:left;padding-left:10px;border:1px solid #e8e8e8',
        reactable(
          data = insurance_sample %>% 
            select(Id) %>%
            mutate(Partition = c(rep("Testing", 3), rep("Training", 7))),
          compact = T,
          bordered = F,
          columns = list(
            Id = colDef(
              name = "Customer ID",
              align = "left"
            ),
            Partition = colDef(
              style = function(value) {
                color <- if (value == "Training") {
                  "#0FA3B199"
                } else {
                  "#f0824899"
                }
                list(background = color)
              }
            )
          )
        )
      )
    )
  ) 
)
```

## Hyperparameter tuning

```{r}
tibble(
  algorithm = rep("Gradient Boosted Trees", 5),
  learning_rate = c(0.001, 0.001, 0.05, 0.01, 0.05),
  ntrees = c(265, 350, 720, 114, 200),
  accuracy = c(0.89, 0.81, 0.76, 0.75, 0.71),
  rank = c(1, 2, 3, 4, 5)
) %>%
  reactable(
    compact = T,
    bordered = T,
    columns = list(
      algorithm = colDef(
        name = "Algorithm",
        align = "left"
      ),
      learning_rate = colDef(
        name = "Learning Rate",
        align = "left"
      ),
      ntrees = colDef(
        name = "Number of Trees",
        align = "left"
      ),
      accuracy = colDef(
        name = "Accuracy",
        align = "left"
      ),
      rank = colDef(
        name = "Rank",
        align = "left"
      )
    )
  ) %>%
  prependContent(
    x = .,
    h3(class = "title", "Example hyperparameter tuning grid showing the hyperparameter combinates sorted from most to least accurate")
  )
```

## Early stopping

### Example of early stopping where training and validation cease once accuracy stops improving {-}

```{r, echo = F}
library(plotly)

min_max_scale <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

errors <- tibble(
  Idx = seq(1,1000,1),
  Error = min_max_scale(sort(rnorm(1000)**2, decreasing = T)),
  Group = c(rep("Actual", 800), rep("Expected", 200))
) 

ggplot(errors, aes(x = Idx, y = Error)) +
  geom_line(aes(color = Group, linetype = Group), lwd = 0.5, show.legend = F) +
  scale_color_manual(values = c("black", "gray50")) +
  annotate("point", x = 800, y = 0.007, shape = 21, size = 5) +
  annotate("text", x = 800, y = 0.19, label = "Early stopping engaged\ndue to no improvements") +
  annotate("segment", xend = 800, x = 800, yend = 0.05, y = 0.11, arrow = arrow(length = unit(0.1, "inches"))) +
  labs(x = "Scoring interval") +
  theme_classic()
```

## Ensembling

```{r}
ensemble_stats <- tibble(
  ID = c(1, 2, 3, 4),
  actuals = c(5.2, 9.5, 1.7, 2.9),
  model1 = c(6.0, 11.1, 2.7, 2.8),
  model2 = c(4.8, 6.9, 1.3, 2.0),
  model3 = c(3.9, 7.9, 2.0, 4.1)
) %>%
  group_by(ID) %>%
  mutate(avg = round(mean(c(model1, model2, model3)), 2))
#mutate(avg_error = avg-actuals)

reactable(
  data = ensemble_stats,
  compact = TRUE,
  bordered = TRUE,
  columns = list(
    ID = colDef(
      name = "Customer ID",
      align = "left"
    ),
    model1 = colDef(
      name = "Model 1",
      align = "left"
    ), 
    model2 = colDef(
      name = "Model 2",
      align = "left"
    ),
    model3 = colDef(
      name = "Model 3",
      align = "left"
    ),
    avg = colDef(
      name = "Ensemble (average of models)",
      align = "left",
      style = function(value) {
        list(background = "#e8e8e8")
      }
    )
  )
) %>%
  prependContent(
    x = .,
    h3(class = "title", "Example of ensembling where predictions from multiple models are combined to increase accuracy")
  )
```

```{r}
ensemble_stats %>%
  ungroup() %>%
  summarize(
    `Model 1` = mean(abs(model1 - actuals)),
    `Model 2` = mean(abs(model2 - actuals)),
    `Model 3` = mean(abs(model3 - actuals)),
    `Ensemble` = mean(abs(avg - actuals))
  ) %>%
  gather(Model, Error) %>%
  arrange(abs(Error)) %>%
  reactable(
    bordered = T,
    compact = T,
  ) %>%
  prependContent(
    x = .,
    h3(class = "title", "Error estimates for individual models and ensemble sorted from lowest to highest error")
  )
```

# Linear regression

### Regression line

```{r}
p <- insurance %>%
  ggplot(aes(x = bmi, y = charges)) +
  geom_point(aes(color = smoker), alpha = 0.1, size = 2) +
  geom_smooth(aes(group = smoker, color = smoker), method = "lm",  se = F, lty = 2) +
  scale_color_manual(values = c("#0FA3B1", "#f08248")) +
  labs(x = "Body mass index",
       y = "Healthcare costs",
       color = "Smoker") +
  theme_classic()

ggplotly(p)
```

# Logistic regression

## Coefficients

```{r}
tibble(
  Feature = c("A", "B", "C", "D"),
  Coefficient = c(3.2, -1.1, 6.5, 0.1)
) %>%
  reactable(
    bordered = T,
    compact = T
  ) %>%
  prependContent(
    x = .,
    h3(class = "title", "Table of feature coefficients")
  )

tibble(
  Dummy = c("Feature coefficients", "Customer actuals"),
  a = c(3.2, 4.1),
  b = c(-1.1, 0.7),
  c = c(6.5, -2.2),
  d = c(0.1, 1.4),
)

tibble(
  customer = rep(1, 4),
  feature = c("A", "B", "C", "D"),
  coeffs = c(3.2, -1.1, 6.5, 0.1),
  actuals = c(4.1, 0.7, -2.2, 1.4)
) %>%
  reactable(
    compact = T,
    bordered = T,
    columns = list(
      customer = colDef(
        name = "Customer ID",
        align = "left"
      ),
      feature = colDef(
        name = "Feature",
        align = "left"
      ),
      coeffs = colDef(
        name = HTML("Coefficient (β)"),
        align = "left"
      ),
      actuals = colDef(
        name = "Actual value (Χ)",
        align = "left"
      )
    )
  )
```

## Equation

## Feature interactions

```{r}

```


# Gradient Boosted Trees

draw.io

